Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-30 demystify.tex
updates

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
% \usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=blue,
}
\usepackage{booktabs}

@@

% \section{Additional Resources}
% \begin{itemize}
%   \item \bibentry{nlpphd}
% \end{itemize}

% \chapter{Additional Resources}

% \chapter{Undergraduate Research Opportunities}

% Prospective PhD students are often unaware of the importance of research experience in their application.  This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
% % This is especially true for students from small colleges where research opportunities are limited.
% This section aims to provide some guidance on how to gain research experience as an undergraduate student. %This section is written together Didier, a PhD student at Mason who did his undergraduate study in Rwanda.

% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.  If you did well or like a class, you can check with the professor of that class for research opportunities.  Sometimes professors would mention that they look for undergraduates on their websites, so do check those out! Even if they say they are looking for graduate students, you still should contact them and ask.  You can also ask your academic adviser or other faculty members for suggestions.  Finally, you can also ask your peers who are already doing research.  You can use the method described in \autoref{sec:contact} to contact professors

% In short, the most important thing is that you need to \emph{ask}!  Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.  Just ask!  The worst thing that can happen is that you get a ``no'' or no response.  But you will never know if you don't ask.  And if you get a ``no'', you can always ask again later.  There are students who got rejected the first time but got accepted the second time.  So don't give up!

% \paragraph{Online Research Communities and Open Source Contributions}:
% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research). Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects. By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.

% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc

% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate. For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles. I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission. I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs. However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions. Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.

% \textbf{Virtual Research Programs}:
% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience. These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers. For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports. Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).

% \textbf{Online Conferences and Workshops}:
% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field. Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations. It is beneficial to also create profiles academic collaboration platforms such as ResearchGate. By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access

% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience. Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
% \usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=blue,
}
\usepackage{booktabs}

@@

\chapter{Visa and all that for International Students}\label{sec:visa}

% \begin{center}
%     \includegraphics[scale=0.8]{files/visa.pdf}
% \end{center}

For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:

\begin{itemize}
\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks). Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.

\item \textbf{Curricular Practical Training (CPT):} This program allows you to participate in internships or practical training that is an integral part of their academic curriculum. CPT must be related to the student's field of study and can be full-time or part-time. It requires prior authorization from the Designated School Official (DSO) and must be completed before graduation.

\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree. For students in STEM fields, there is an additional 24-month extension available. OPT requires prior application and approval from the United States Citizenship and Immigration Services (USCIS).

\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year. This means taking a minimum number of credits each semester, as defined by your program. Dropping below full-time status can result in loss of visa status.
\end{itemize}

\includepdf[pages=-]{files/visa.pdf}

\section{For Spouses and Children}

The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:

For Spouse: 
\begin{itemize}
\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
\end{itemize}
For Children:
\begin{itemize}
\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
\end{itemize}
F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
